---
title: "MC-D607 Final Project HUD-API"
author: "Marco Castro"
date: "2024-11-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(jsonlite)
library(tibble)
library(httr2)
library(stringr)
library(rvest)
library(textutils)

# Read in our API global key
source("credentials.R")
```

## Fair Market Rents (FMR) vs Market Rate Rents

This project aims to ascertain whether the market rate rental costs for 1 and 2 bedroom apartments are consistent with the federal government's [Fair Market Rents (FMR)](https://www.huduser.gov/periodicals/ushmc/winter98/summary-2.html) — the 40th percentile of gross rents for typical rental units paid by tenants that moved within the last 20 months. FMR is used to calculate benefits such as the Housing Choice Vouchers used to help unhoused individuals find permanent housing. 

As an annually-released estimate, FMR data does not capture the real-time economic realities that drive local housing markets. This means that some individuals that are approved for housing assistance benefits are unable to find an apartment locally that they can pay for with their approved voucher amount based on their area's FMR. In other words, market rates tend to outpace FMR estimates when market conditions quickly drive rental prices up in certain areas. 

This project uses FMR data pulled from HUD using their publicly available [API](https://www.huduser.gov/portal/dataset/fmr-api.html). Market rate data was scraped from the rental listing site [Trulia](http://trulia.com) in three distinct metropolitan areas: Atlanta, GA, Buffalo, NY and San Diego, CA to obtain an up-to-date snapshot of real rental costs that can be compared against data from HUD's FMR Dataset. The comparison will help to determine if the FMR data released earlier this year for fiscal year Oct 2024-Sept 2025 is still relevant to current housing costs in these areas.

```{r declare-cities}
# A1: declare the cities we will analyze 
target_cities <- c('Atlanta-Sandy Springs-Roswell','Buffalo-Cheektowaga-Niagara Falls','San Diego-Carlsbad')
```

## A: Getting FMR data using the HUD API

The section below declares the global variables and functions for to call the [HUD FMR & IL API](https://www.huduser.gov/portal/dataset/fmr-api.html) using the `httr2` package using a HUD API key (A2). 

```{r get_hud_data_with_httr2}

# A2: Custom function to call HUD FMR & IL API
call_hud <- function(endpoint) {
  # declare constants
  domain <- "https://www.huduser.gov"
  method   <- "fmr" # alt method is il or mtspil
  path <-paste("hudapi/public/", method, "/", endpoint,sep="")
  
  # init request
  req <- request(domain) |> 
    req_headers("Accept" = "application/json") |> 
    req_auth_bearer_token(token) |>
    req_url_path(path) |>
    req_user_agent("Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/64.0.3282.140 Safari/537.36 Edge/18.17763")
  
  # call api
  resp <- req_perform(req)
  
  # parse response
  json_resp <- resp |> 
    resp_body_string() |>
    fromJSON() 
  
  # return parsed response
  return(json_resp) 
}

```

### Retrieving the Metro Ids

The custom function `call_hud` is used in the section below (A3) to request the HUD metropolitan area id (`entity_id`) for each of our target cities. This id will then be used to retrieve the FMR for each of the zipcodes that make up the area as well as the estimate for the area overall. The request returns a json obj for all metropolitan areas. After converting the response to a dataframe (A4), we will need to separate the __area_name__ column to cities, states, and area type. Finally, we filter for using the target cities array declared earier.

```{r get-metro-area-ids}
  # A3: Fetch Metro Area Ids from API
  metros <- call_hud("listMetroAreas")

  # A4: cleanup and filter list of metro areas 
  #     for target areas
  as.data.frame(metros)
  metro_areas <- as.data.frame(metros) |>
    # split area_name col into city and type
    separate_wider_regex(
      area_name, c(cities= ".*", ",\\s", type=".*")
    ) |>
    # split type col into state and type
    mutate(
      state = substring(type,1,2),
    ) |>
    # filter by our target areas
    filter(cities %in% target_cities) |>
    select(c('cbsa_code', 'cities', 'state')) 

  
glimpse(metro_areas)  
```

### Retrieving FMRs for Target Cities

Now that we have the ids for the metropolitan areas, we can call the `data` method of HUD's API to retrieve the FMR values for each metropolitan area and its corresponding zip codes (A5). I created a custom function `get_metro_data` that requests the FMR estimates the API and clean up the response. This function is called from a loop that iterates through each of our target areas. Finally, the loop uses `rbind` to merge the dataframes for each target city into a single dataframe

```{r get-entity-data}

# A5: Call, clean-up and merge metro area data
# Custom function to call and clean up metro area data
get_metro_data <- function(row) {
  # use "cbsa_code" col as entity_id for API call
  entity_resp <- call_hud(paste("data", row$cbsa_code, sep="/"))
  print(entity_resp)
  # clean up data
  df <- as.tibble(entity_resp$data$basicdata) |>
    rename_all(~tolower(str_trim(str_replace_all(., "-", "_")))) |>
    rename(studio = efficiency) |>
    mutate( 
      cbsa_code = row$cbsa_code
    ) |>
    relocate(cbsa_code, .before=zip_code) 
}

# init df with first row
metro_fmrs <- get_metro_data(metro_areas[1,])

# loop through all other metro areas
for (i in 2:nrow(metro_areas)) {
  df <- get_metro_data(metro_areas[i,])
  metro_fmrs <- rbind(metro_fmrs, df)
}

head(metro_fmrs)

zip_codes <- metro_fmrs |>
  filter(zip_code != 'MSA level') |>
  select(c("cbsa_code", "zip_code"))
```


# Tidying the FMR data

In section A6, I used `pivot_longer` to break up each FMR cost estimate into its own row broken up by apartment type in each apartment. I then used `mutate` to convert the bedroom classifications to an integer representation based on the number of bedrooms. For example, a studio was given a value of 0, `one_bedroom` was given a value of 1, etc. 

```{r tidy-fmr-data}
# A6:  tidy data
metro_fmr_clean <- metro_fmrs |>
  pivot_longer(
    cols = studio:four_bedroom,
    names_to = c("bedrooms"),
    values_to = "fmr"
  ) |>
  mutate(
    bedrooms = case_when(
      bedrooms == 'studio' ~ as.integer(0),
      bedrooms == 'one_bedroom' ~ as.integer(1),
      bedrooms == 'two_bedroom' ~ as.integer(2),
      bedrooms == 'three_bedroom' ~ as.integer(3),
      bedrooms == 'four_bedroom' ~ as.integer(4)
    )
  ) |>
  arrange(zip_code)

head(metro_fmr_clean, n=10)

```

# B: Scraping Market Rate Data

This script scrapes the [Trulia] ("https://www.trulia.com/").

```{r init-scraper}

# B1: Declare parser constants

# init market rate tibble
#real_estate_df <- tibble(
#  zip_code = character(),
#  bedrooms = integer(),
#  market_rate = integer()
#)

# custom function to remove misc chars
cleanup_price <- function(s) {
  t <- str_replace_all(s, "([$,])", "") #remove dollar signs and commas
  x <- gsub("^([0-9]+).*", "\\1", t) #only return first group of integers
  print(x)
  return(as.integer(x))
} 

# global vars
domain <- "https://www.trulia.com"

# update user agents
user_agent <- "Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/64.0.3282.140 Safari/537.36 Edge/18.17763"
set_config(add_headers(`User-Agent` = user_agent))


```

```{r}


zip_codes <- zip_codes |>
  filter(cbsa_code == 'METRO12060M12060')
```

```{r scrape-trulia}

delay <- 103
# custom function for delayed scraping
throttled_GET <- slowly(
  ~read_html(.),
  rate = rate_delay(delay)
)

# C2: Scrape Real Estate Data



# can be altered for multipage scraping
beds <- 0
current_page <- 1
start <- 1
end <- nrow(zip_codes)

# loop through zip codes list
for (i in start:end) {
  # get zip code at current index
  zip <- zip_codes[i,]$zip_code
  # randomly set the delay in seconds
  delay <- 45 + round(runif(1) * 25)
  
  print(delay)
  # print current zip
  print(zip)
  
  # build our path based on the current zip, beds and page vars 
  path <- paste(domain, "/for_rent/", zip, "_zip/", beds, "_beds/", current_page, "_p/", sep="")
  
  # send path to our throttled function 
  page_reponse <- throttled_GET(path)
  
  # find price from current page html
  result_container <- page_reponse %>% 
    html_element(xpath='//ul[@data-testid="search-result-list-container"]' )
    listings <- result_container %>% 
    html_elements(xpath = '//li//div[@data-testid="property-price"]' ) %>% 
    html_text()  

  # C3: Append to parsed real estate dataframe
  # loop through all listing prices 
  # and append to market rate tibble
  for (s in listings)
    real_estate_df <- real_estate_df |>
      add_row(zip_code = zip, bedrooms = beds, market_rate = cleanup_price(s))
  
  
  write_csv(real_estate_df, "trulia_data.csv")
}


head(real_estate_df, n=10)
```



```{r dump_parsed_data }
write_csv(real_estate_df, "trulia_data.csv")
```






